On Sample Size Needed for Block Bootstrap Confidence Intervals to Have Desired Coverage Rates

### 2024-01-02 Manuscript/manuscript.tex
initial feedback

\citet{lahiri1999theoretical} finds that moving block bootstrap has better 
performance than non-overlapping block bootstrap. Additionally, moving block 
bootstrap with nonrandom
block sizes results in lower mean-squared errors than moving block bootstrap
with random block sizes. \citet{buhlmann1999block} notes that a drawback of 
block bootstrap is that it heavily depends on block size, which has to be chosen
by the user of the method.
Even when using the appropriate settings, as noted by
\citet{buhlmann2002bootstraps} notes some general drawbacks of block 
bootstrap --- with respect to how reasonably it imitates the data-generating 
process. In addition, although block bootstrap is primarily used for stationary 
time series, it can be outperformed by other bootstrap schemes for linear time
series and categorical processes.

@@

\paragraph{Recentered Percentile CI}
We propose a CI that is centered at the original point estimate and uses the 
variation in the bootstrap estimates to construct the error bound. This interval 
requires
the computation of $\bar\theta_n^B = n^{-1}\sum_{b=1}^B \hat\theta_n^{(b)}$, the 
mean of all bootstrap point estimates. A $(1 - \alpha)100\%$ CI is
centered around $\hat\theta_n$ and can be written as 
\[
(\hat\theta_n + \hat\theta_{n, \alpha/2}^{B} - \bar\theta_n^{B}, \quad
\hat\theta_n + \hat\theta_{n, 1 - \alpha/2}^{B} - \bar\theta_n^{B}).
\]
It is not necessarily symmetric, as different critical values are used to 
compute the lower and upper bounds. It has the same width as the percentile CI.

@@

\subsection*{Design}
We designed a simulation study to compare the performance of the different block 
bootstrap CI methods. In particular, we generated time series $X_t$ from a 1st 
order autoregressive (AR(1)) process:
\begin{equation*}
X_t = \phi X_{t-1} + \epsilon_t,
\end{equation*}
where $\phi$ is an autoregressive coefficient, and $\epsilon_t$ is a series of
independent errors from a normal distribution with mean zero and variance
$\sigma_{\epsilon}^2$. The strength of the serial dependence is controlled by
$\phi$, which was set to five levels: $\{-0.4, -0.2, 0.0, 0.2, 0.4\}$. We only
used serial dependences as strong as 0.4, because we only seek to 
establish the 
general trend as the strength of the autocorrelation 
increases, and how it varies depending on the sign of the autocorrelation and 
the parameter of interest. The 
series $X_t$ has mean zero and variance 
$\sigma_x^2 = (1 - \phi^2) \sigma_{\epsilon}^2$, so for each value of $\phi$, we 
set $\sigma_{\epsilon}^2 = 1 / (1 - \phi^2)$ such that $\sigma_x^2 = 1$.

@@

Additionally, we evaluated the performance of block bootstrap when used to 
estimate
the mean, standard deviation, and the lag-1 autocorrelation coefficient of a
stationary series with an exponential marginal distribution. The motivation of
for this study was to investigate if the results are robust to nonnormal
marginal distributions.

@@

For estimating the mean parameter~$\mu$, \textbf{Figure~\ref{fig:mu1}} suggests 
that all methods eventually approach correct coverage of $\mu$ as sample size 
increases. Student's $t$ CIs appear to need the smallest sample size to achieve 
correct coverage, except for samples with strong negative dependence, in which 
case, they actually over-cover $\mu$ for smaller sample sizes. For instance, for 
a sample with $n = 100$ and $\phi = -0.4$, the lower bound for a Students $t$ 
CI's coverage of $\mu$ is greater than 95\%, whereas the coverage intervals for 
other methods contain 95\%. The standard normal, percentile, BC, and BCA, and 
recentered percentile CIs require similar sample sizes to recover $\mu$ at the 
nominal level for all combinations of $n$ and $\phi$. All methods seem to 
require a smaller sample to recover $\mu$ at the nominal rate when dealing with 
negative dependence versus positive dependence. For example, BC CIs recover 
$\mu$ for $n \geq 100$ when $\phi = -0.2$, but they only recover $\mu$ for 
$n \geq 800$ when $\phi = 0.2$. In addition, as a negative dependence gets 
stronger, holding everything else equal, coverage increases, which lead to the 
Student $t$ CI's aforementioned over-coverage. As a positive dependence gets 
stronger, holding everything else equal, coverage decreases, and a larger sample 
is necessary to recover~$\mu$. A possible explanation for this is that if a 
stationary series has a positive 
autocorrelation, the effective sample
size is decreased, whereas is a series has a negative autocorrelation, the
effective sample size is increased. Additionally, this seems to have a 
greater effect on the the estimation of the location parameter versus that
of the scale parameter or temporal dependence parameter.

@@

For an exponential marginal distribution, the empirical coverage rates for 
$\mu$, $\sigma_w$, %\eds{define $\sigma_w$} 
%\mc{added definitions for $\mu$ and $\sigma_w$ in the above paragraph} 
and 
the lag-1 autocorrelation coefficient, %\mc{replaced $\phi$ here, as I understand
%it, we are not using $\phi$ in reference to the temporal dependence of the 
%exponentially distributed series?}, 
as well as 95\%
confidence intervals of the real coverage are displayed in 
\textbf{Figures~\ref{fig:exp_mu1}--\ref{fig:exp_phi2}}. 
Additionally, a set of 50 CIs are displayed for each of the percentile, 
BC, and recentered percentile approaches for exponentially distributed
samples of $n \in \{800, 1600\}$ with lag-1 autocorrelation coefficient 0.355
($\phi = 0.4$) in Figure~\ref{fig:eri}.

>>>>

\citet{lahiri1999theoretical} finds that moving block bootstrap has better 
performance than non-overlapping block bootstrap. Additionally, moving block 
bootstrap with nonrandom
block sizes results in lower mean-squared errors than moving block bootstrap
with random block sizes. \citet{buhlmann1999block} notes that a drawback of 
block bootstrap is that it heavily depends on block size, which has to be chosen
by the user of the method.
Even when using the appropriate settings, as noted by
\citet{buhlmann2002bootstraps} notes some general drawbacks of block 
bootstrap --- with respect to how reasonably it imitates the data-generating 
process. In addition, although block bootstrap is primarily used for stationary 
time series, it can be outperformed by other bootstrap schemes for linear time
series and categorical processes.
\eds{If we include all these drawbacks here, we need to also provide some good 
features of block bootstrap as well, otherwise, why do we care about their CI 
performance?}

@@

\paragraph{Recentered Percentile CI}
We propose a CI that is centered at the original point estimate and uses the 
variation in the bootstrap estimates to construct the error bound. 
\eds{The motivation behind proposing such an interval was based on the 
simulation performance of the BC and BCA intervals, which will be discussed
further in the Results section.} 
This interval 
requires
the computation of $\bar\theta_n^B = n^{-1}\sum_{b=1}^B \hat\theta_n^{(b)}$, the 
mean of all bootstrap point estimates. A $(1 - \alpha)100\%$ CI is
centered around $\hat\theta_n$ and can be written as 
\[
(\hat\theta_n + \hat\theta_{n, \alpha/2}^{B} - \bar\theta_n^{B}, \quad
\hat\theta_n + \hat\theta_{n, 1 - \alpha/2}^{B} - \bar\theta_n^{B}).
\]
It is not necessarily symmetric, as different critical values are used to 
compute the lower and upper bounds. It has the same width as the percentile CI.

@@

\subsection*{Design}

\eds{We compared the performance of the different block bootstrap CI methods 
under two marginal distributions: normal and exponential.}

\paragraph{Normal Marginal Distribution}
%We designed a simulation study to compare the performance of the different block 
%bootstrap CI methods. In particular, 
We generated time series $X_t$ from a 1st 
order autoregressive (AR(1)) process:
\begin{equation*}
X_t = \phi X_{t-1} + \epsilon_t,
\end{equation*}
where $\phi$ is an autoregressive coefficient, and $\epsilon_t$ is a series of
independent errors from a normal distribution with mean zero and variance
$\sigma_{\epsilon}^2$. The strength of the serial dependence is controlled by
$\phi$, which was set to five levels: $\{-0.4, -0.2, 0.0, 0.2, 0.4\}$. We only
used serial dependences as strong as 0.4, because we only seek to 
establish the 
general trend as the strength of the autocorrelation 
increases, and how it varies depending on the sign of the autocorrelation and 
the parameter of interest. The 
series $X_t$ has mean zero and variance 
$\sigma_x^2 = (1 - \phi^2) \sigma_{\epsilon}^2$, so for each value of $\phi$, we 
set $\sigma_{\epsilon}^2 = 1 / (1 - \phi^2)$ such that $\sigma_x^2 = 1$.

@@

\paragraph{Exponential Marginal Distribution}
Additionally, we evaluated the performance of block bootstrap when used to 
estimate
the mean, standard deviation, and the lag-1 autocorrelation coefficient of a
stationary series with an exponential marginal distribution. The motivation of
for this study was to investigate if the results are robust to nonnormal
marginal distributions.

@@

\paragraph{Normal Marginal Distribution}
For estimating the mean parameter~$\mu$, \textbf{Figure~\ref{fig:mu1}} suggests 
that all methods eventually approach correct coverage of $\mu$ as sample size 
increases. Student's $t$ CIs appear to need the smallest sample size to achieve 
correct coverage, except for samples with strong negative dependence, in which 
case, they actually over-cover $\mu$ for smaller sample sizes. For instance, for 
a sample with $n = 100$ and $\phi = -0.4$, the lower bound for a Students $t$ 
CI's coverage of $\mu$ is greater than 95\%, whereas the coverage intervals for 
other methods contain 95\%. The standard normal, percentile, BC, and BCA, and 
recentered percentile CIs require similar sample sizes to recover $\mu$ at the 
nominal level for all combinations of $n$ and $\phi$. All methods seem to 
require a smaller sample to recover $\mu$ at the nominal rate when dealing with 
negative dependence versus positive dependence. For example, BC CIs recover 
$\mu$ for $n \geq 100$ when $\phi = -0.2$, but they only recover $\mu$ for 
$n \geq 800$ when $\phi = 0.2$. In addition, as a negative dependence gets 
stronger, holding everything else equal, coverage increases, which lead to the 
Student $t$ CI's aforementioned over-coverage. As a positive dependence gets 
stronger, holding everything else equal, coverage decreases, and a larger sample 
is necessary to recover~$\mu$. A possible explanation for this is that if a 
stationary series has a positive 
autocorrelation, the effective sample
size is decreased, whereas is a series has a negative autocorrelation, the
effective sample size is increased. Additionally, this seems to have a 
greater effect on the the estimation of the location parameter versus that
of the scale parameter or temporal dependence parameter.

@@

\paragraph{Exponential Marginal Distribution}
For an exponential marginal distribution, the empirical coverage rates for 
$\mu$, $\sigma_w$, %\eds{define $\sigma_w$} 
%\mc{added definitions for $\mu$ and $\sigma_w$ in the above paragraph} 
and 
the lag-1 autocorrelation coefficient, %\mc{replaced $\phi$ here, as I understand
%it, we are not using $\phi$ in reference to the temporal dependence of the 
%exponentially distributed series?}, 
as well as 95\%
confidence intervals of the real coverage are displayed in 
\textbf{Figures~\ref{fig:exp_mu1}--\ref{fig:exp_phi2}}. 
Additionally, a set of 50 CIs are displayed for each of the percentile, 
BC, and recentered percentile approaches for exponentially distributed
samples of $n \in \{800, 1600\}$ with lag-1 autocorrelation coefficient 0.355
($\phi = 0.4$) in Figure~\ref{fig:eri}.

### 2024-01-02 Manuscript/reply.tex
initial feedback

@@

@@

The formula for the "parametric" CIs concerns the distribution of the 
bootstrapped statistic, but we actually expect such CIs to work for different
error structures.

@@

This terminology was used because, \citet{efron1993introduction} refers to such 
CIs as 
"confidence intervals based on 
bootstrap 'tables'." We have changed these sentences as shown below:

@@

We have added this comment to the discussion:

@@

@@

Because of the deterioration for $\phi$, we found it hard to fix the limits on 
the 
y-axis without hurting the visualization for the other rows. Below is an 
example:

@@

The original sentence has been replaced with:

@@

@@

@@

We have included this comment in the Design subsection:

@@

These intervals are discussed in \citep{efron1993introduction}. A citation
is now provided in the last paragraph.

@@

We have included this discussion of problems with block bootstrap in the
introduction:

\citet{lahiri1999theoretical} finds that moving block bootstrap has better 
performance than non-overlapping block bootstrap. Additionally, moving block 
bootstrap with nonrandom
block sizes results in lower mean-squared errors than moving block bootstrap
with random block sizes. \citet{buhlmann1999block} notes that a drawback of 
block bootstrap is that it heavily depends on block size, which has to be chosen
by the user of the method.
Even when using the appropriate settings, as noted by
\citet{buhlmann2002bootstraps} notes some general drawbacks of block 
bootstrap --- with respect to how reasonably it imitates the data-generating 
process. In addition, although block bootstrap is primarily used for stationary 
time series, it can be outperformed by other bootstrap schemes for linear time
series and categorical processes.

>>>>

\eds{Also added comment in the Recentered Percentile CI paragraph in the 
BLOCK BOOTSTRAP CIS section. Please check and add a comment here if you like it.}

@@

\eds{I also added subsubsections (paragraphs) for the Normal and Exponential 
Marginal Distributions under Design and Results subsections.
Is it possible to put to make the [Simulation Study] Design and 
[Simulation Study] Results as two separate 
sections instead of placing each as subsections of the Simulation Study section,
or does the journal formatting prohibit this?  If you can change this, then the 
Normal and Exponential Marginal subsubsections can be updated to subsections
under the Simulation Study Design and Simulation Study Results sections.}

@@

The formula for the "parametric" CIs concerns the distribution of the 
bootstrapped statistic, but we actually expect such CIs to work for different
error structures. \eds{you will need to add more here.  This concerns the
(asymptotic) distribution of the statistics, not the underlying error 
structure.}

@@

Thank you for the comment.  This terminology was used because 
\citet{efron1993introduction} refers to such CIs as 
"confidence intervals based on 
bootstrap 'tables'." We have changed these sentences as shown below:

@@

We agree. We have added this comment to the discussion:

@@

\eds{I don't think this answers the question.  The reviewer is saying that 0.95 
is close to the boundary of 1, so the undercoverage observed intervals
(especially when phi -0.4) may be due, at least in part, to the flaws 
of the asymptotic Wald interval you are using. 
\@ Jun: do you think he should try exact Clopper-Pearson
intervals for at least a few cases of undercoverage to verify it is not the Wald
interval that is causing the undercoverage, then comment on this in at least the
reply (and possibly also in the main paper)?}

@@

Because of the deterioration for $\phi$, we found it hard to fix the limits on 
the 
y-axis without hurting the visualization for the other rows. Below is an 
example:  \eds{refer to the Figure label, and avoid using `below' and `above'
terminology with latex because the figures are not always positioned as we may
have intended.}

@@

Thank you for pointing this out. The original sentence has been replaced 
with:

@@

\eds{I only saw this citation in the abstract, not the discussion.  
Please remove this citation from the abstract and place in the discussion 
section where the sentence in question appears. In general,
try to avoid placing any citations in the abstract.}

@@

\eds{\@ Jun: what do you think about putting the $l = \lceil 2n^{1/3} \rceil$
results in the Appendix, but keeping the conclusions regarding how the 
$l = \lceil n^{1/3} \rceil$ and $l = \lceil 2n^{1/3} \rceil$ simulations differ
in the main text?}

@@

Thank you for the comment. We have now included this comment in the Design 
subsection:

@@

We apologize for the omission. These intervals are discussed in 
\citep{efron1993introduction}. A citation is now provided in the last paragraph.

@@

Thank you for the comment. We have included this discussion of problems 
with block bootstrap in the introduction:

\citet{lahiri1999theoretical} finds that moving block bootstrap has better 
performance than non-overlapping block bootstrap. Additionally, moving block 
bootstrap with nonrandom
block sizes results in lower mean-squared errors than moving block bootstrap
with random block sizes. \citet{buhlmann1999block} notes that a drawback of 
block bootstrap is that it heavily depends on block size, which has to be chosen
by the user of the method.
Even when using the appropriate settings, 
\citet{buhlmann2002bootstraps} notes some general drawbacks of block 
bootstrap --- with respect to how reasonably it imitates the data-generating 
process. In addition, although block bootstrap is primarily used for stationary 
time series, it can be outperformed by other bootstrap schemes for linear time
series and categorical processes.

\eds{If we include all these drawbacks in the intro, we need to also provide 
some good features of block bootstrap as well, otherwise, why do we care about 
their CI performance?}
